Software Architectures for Integrating Workflow Engines in Science Gateways

### 2016-05-13 paper.tex
Some fixes following @SilviaOlabarriaga's comments

\note{Sil}{i changed above a few occurrences of "diagrammed". i know it is correct, but i am not used to seeing it. this is a matter of style, just forget my changes below then}

@@

The interactions involved in the architectures are described below and
labeled (a,b,c, ...) as in Figure~\ref{fig:architectures}.

@@

  another. This interaction may not be available or possible
  for every workflow language. It has been developed mostly for
  translation between well-structured and relatively simple workflow languages such as GWorkflowDL
  and Scufl~\cite{OLAB-09}, and the 4 languages that were involved in the
  SHIWA initiative.
 \note{Silvia}{Should we add here the 4?
 	and for translation among the 5 workflow systems in the SHIWA project: Askalon, MOTEUR, Taverna, Triana and WS-PGRADE)} 

@@

See Figure~\subref*{archi:service}. The workflow engine is available
externally to the science gateway, as a service. The science gateway
controls the service through a specific interaction (\texttt{d}) that
might be implemented as a web-service call (e.g., RESTful or SOAP), as
a command-line or as any other method that offers a well-defined
interface to the workflow engine. The workflow engine might be invoked
either as a black box that completely masks the infrastructure and
tasks, or as a white box that allows for some interaction with
them. The workflow engine is responsible for controlling the tasks on
the infrastructure (\texttt{b}) and for performing the required data
transfers to execute them (\texttt{c$_2$}). User data is usually
managed through the science gateway (\texttt{c$_1$}), although it
might as well be delivered by the workflow engine directly to the
user.
\note{Sil}{about c1: should it also show from engine to infra in the figure? this is what i understand from the above sentence}

@@

The science gateway also transfers both user and workflow data across
the infrastructure, so that interactions \texttt{c$_1$} and
\texttt{c$_2$} are both covered by \texttt{c}. 
\note{Sil}{is this next sentence relevant?}In practice, both
\texttt{c$_1$} and \texttt{c$_2$} can be implemented using the same
pieces of code.

@@

Differences in other specific features should not be neglected. Nested
workflows (with service invocation) and workflow import (with service
invocation) are the only architectures that support meta-workflows;
task encapsulation and nested workflows may complexify
fine-grained debugging. \closednote{Rafael}{I would suggest to provide
  some more examples when this could be interesting.}{Tristan}{Done.}
The availability of fine-grained debugging information may also be
critical to efficient user support.

@@

A few limitations should be considered when using the results of our
evaluation. First, our evaluation methods aimed to provide comparative
metrics, however the high-level of abstraction used to derive these
metrics hinders the complexity of real systems to some
extent.\closednote{Naj}{Here, you were bashing your article harshly,
  so I rewrote it}{Tristan}{yup, thanks!}. In particular, the
presented architectures are abstract patterns that may be mixed
together in actual systems.\answerednote{Naj}{something about "theoretical
  pattern", and "blending" is jarring, are these common words in the
  computer sci context?}{Tristan}{Better now?} The distinction between tight integration and
service invocation, for instance, may not always be that clear in
practice. Service invocation may also be combined with task
encapsulation in some cases. However, the criteria discussed in this
work would still apply to broadly compare and categorize such cases of
hybrid architectures.

>>>>

\closednote{Sil}{i changed above a few occurrences of "diagrammed". i know it is correct, but i am not used to seeing it. this is a matter of style, just forget my changes below then}{Tristan}{That's perfect, thanks!}

@@

The interactions involved in the architectures are described below and
labeled (\texttt{a},\texttt{b},\texttt{c}, \ldots) as in Figure~\ref{fig:architectures}.

@@

  another. This interaction may not be available or possible for every
  workflow language. It has been developed mostly for translation
  between well-structured and relatively simple workflow languages
  such as GWorkflowDL and Scufl~\cite{OLAB-09}, and for translation
  among the 5 workflow systems in the SHIWA project: Askalon, MOTEUR,
  Taverna, Triana and WS-PGRADE.  \closednote{Silvia}{Should we add here the
    4?  and for translation among the 5 workflow systems in the SHIWA
    project: Askalon, MOTEUR, Taverna, Triana and WS-PGRADE)}{Tristan}{Yes, done!}

@@

See Figure~\subref*{archi:service}. The workflow engine is available
externally to the science gateway, as a service. The science gateway
controls the service through a specific interaction (\texttt{d}) that
might be implemented as a web-service call (e.g., RESTful or SOAP), as
a command-line or as any other method that offers a well-defined
interface to the workflow engine. The workflow engine might be invoked
either as a black box that completely masks the infrastructure and
tasks, or as a white box that allows for some interaction with
them. The workflow engine is responsible for controlling the tasks on
the infrastructure (\texttt{b}) and for performing the required data
transfers to execute them (\texttt{c$_2$}). User data is usually
managed through the science gateway (\texttt{c$_1$}), although in some
variations of the architecture (not represented in
Figure~\ref{fig:architectures}) it might as well be delivered by the
workflow engine directly to the user. \answerednote{Sil}{about c1: should it
  also show from engine to infra in the figure? this is what i
  understand from the above sentence}{Tristan}{Actually c1 is either
  coming from the science gatewray or from the workflow engine but
  then it merges with c2. I don't think we can cover both cases in a
  single architecture, the latter case would be another variation of
  this architecture. I modified a bit the text, is it clearer now?}

@@

The science gateway also transfers both user and workflow data across
the infrastructure, so that interactions \texttt{c$_1$} and
\texttt{c$_2$} are both covered by \texttt{c}. 
\closednote{Sil}{is this next sentence relevant?}{Tristan}{I just wanted to  justify why these two interactions can be grouped but I guess you're right (I removed the sentence.}%In practice, both
%\texttt{c$_1$} and \texttt{c$_2$} can be implemented using the same
%pieces of code.

@@

Differences in other specific features should not be neglected. Nested
workflows (with service invocation) and workflow import (with service
invocation) are the only architectures that support
meta-workflows. Besides, task encapsulation and nested workflows may
complexify fine-grained debugging. \closednote{Rafael}{I would suggest
  to provide some more examples when this could be
  interesting.}{Tristan}{Done.}  The availability of fine-grained
debugging information may also be critical to efficient user support.

@@

A few limitations should be considered when using the results of our
evaluation. First, our evaluation methods aimed to provide comparative
metrics, however the high-level of abstraction used to derive these
metrics hinders the complexity of real systems to some
extent.\closednote{Naj}{Here, you were bashing your article harshly,
  so I rewrote it}{Tristan}{yup, thanks!}. In particular, the
presented architectures are abstract patterns that may be mixed
together in actual systems.\closednote{Naj}{something about "theoretical
  pattern", and "blending" is jarring, are these common words in the
  computer sci context?}{Tristan}{Better now?} The distinction between tight integration and
service invocation, for instance, may not always be that clear in
practice. Service invocation may also be combined with task
encapsulation in some cases. However, the criteria discussed in this
work would still apply to broadly compare and categorize such cases of
hybrid architectures.
